\pkg{Rgbp}: Bayesian Hierarchical Modeling and Frequentist Method Check

### 2013-04-14 jss/article.tex
some corrections

Considering the type of data, it makes sense to assume the number of deaths in each hospital has independent Poisson distribution given the unknown true rate parameter $\lambda_{j}$: $z_{j}\vert \lambda_{j}\stackrel{ind}{\sim} \textrm{Poisson}(n_{j}\lambda_{j})$, $j=1, \ldots, 31,$ where $n_{j}$ can be interpreted as an exposure (not necessarily an interger). This dataset is also included in the package and can be called by `\code{R> data(hospital)}' on \proglang{R}.

@@

where $\log(\frac{p_{0j}}{1-p_{0j}}) =x_{j}'\beta~\textrm{and}~j=1, \ldots, 18$. Then posterior distribution is
\begin{equation}
p_{j}\vert \textbf{z}, \beta, r \stackrel{ind}{\sim}\textrm{Beta}(rp_{0j}+n_{j}y_{j},~r(1-p_{0j})+n_{j}(1-y_{j})\sim\textrm{Beta}\bigg[p^{\ast}_{j},~ \frac{p^{\ast}_{j}(1-p^{\ast}_{j})}{r+n_{j}+1}\bigg],
\end{equation}
where $p^{\ast}_{j}\equiv(1-B_{j})y_{j}+B_{j}p_{0j}$, $B_{j}\equiv\frac{r}{r+n_{j}}$, and $y_{j}\equiv\frac{z_{j}}{n_{j}}$, $j=1,\ldots,18$.

@@

where $m$ is the number of regression coefficients. For $\beta$, it is reasonable choice to take flat (non-informative) distribution because information about the location gets plentiful as the number of groups increases. Another flat prior on the second-level variance component  $A$ (or $1/r$) that we suggest here will guarantee a posterior propriety under the moderate conditions and will bring a good repeated sampling property to us.
\\

@@

Once we estimate these two moments of shrinkage, we can also estimate the posterior moments given only data. Taking the Normal model as an example,  $E(\mu_{j}\vert \textbf{y})=E(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})$ by Adam's law and $Var(\mu_{j}\vert \textbf{y})=E(Var(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})+Var(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})$ by Eve's law. Note that both $E(\mu_{j}\vert \textbf{y}, r, A)$ and $Var(\mu_{j}\vert \textbf{y}, r, A)$ are a function of shrinkage in (3) and Adam's and Eve's laws are taking conditional expectation and variance over it given data.

@@

\subsection{Pseudo-data Generation Process}
Figure 1 will be helpful to understand this process. As we can see in (8), the distribution of each true batting average ($p_{j},~j=1,\ldots, 18$) depends on two hyper-parameters, $r$ and $\beta_{(m\times1)}$, where $m$ is the number of regression coefficients. So, once we fix these hyper-parameters at specific values, we can generate true batting averages. Suppose we sampled 500 $\mathbf{p}_{(18\times1)}$'s, \emph{i.e.}, $\{\mathbf{p}^{(i)}_{(18\times1)},~i=1, \ldots, 500\}$ from the prior distribution in (8), where $r$ and $\beta$ are given. Then, we can also generate $\{\mathbf{z}^{(i)}_{(18\times1)},~i=1, \ldots, 500\}$ given each $p_{ij},$ where $i$ indicates $i$-th pseudo-dataset and $j$ does $j$-th player. Next, \code{coverage} fits the Binomial hierarchical model 500 times on $\{(\mathbf{z}^{(i)}_{(18\times1)}, \mathbf{n}_{(18 \times 1)}),~i=1, \ldots, 500\}$ to obtain 500$\times$18 interval estimates.  After this generating process, as for the first player, we will have 500 pairs of ($z_{i1}$, $p_{i1}$) and 500 interval estimates, ($\hat{p}_{i1, low}, \hat{p}_{i1, upp}$), $i=1, \ldots, 500$.
\begin{figure}[h]
\begin{center}
\includegraphics[width=5cm]{process.png}
\caption{Pseudo-data generating process}
\end{center}
\end{figure}

@@

Assuming an additional hierarchy is reasonable, a model-fitting process begins.  Since the true death rate per exposure after CABG surgery might be small and the caseloads ($n_{j}$) look relatively much bigger than the number of deaths ($z_{j}$), the Poisson distribution would be our first choice to describe the uncertainty in our data. Next, \code{gbp} will help us fit the Poisson multi-level model with the Gamma conjugate prior distribution on the true death rate $\lambda_{j}$ whose mean is 0.02 ($\lambda_{0}=0.02$) as described in 3.2 . For reference, the number of regression coefficients ($m$) is 0 because we do not need to estimate the prior mean via any regression.
\begin{CodeChunk}
\begin{CodeInput}
R> p <- gbp(z, n, mean.PriorDist = 0.02, model = "pr")
R> p
\end{CodeInput}
\begin{CodeOutput}
Summary for whole observations:

         obs.mean    n prior.mean shrinkage low.intv post.mean upp.intv post.sd
1           0.045   67       0.02     0.834    0.011     0.024    0.042   0.008
2           0.029   68       0.02     0.832    0.010     0.022    0.038   0.007
3           0.024  210       0.02     0.616    0.011     0.021    0.035   0.006
4           0.043  256       0.02     0.568    0.017     0.030    0.046   0.008
5           0.033  269       0.02     0.556    0.015     0.026    0.041   0.007
6           0.044  274       0.02     0.551    0.018     0.031    0.047   0.008
7           0.043  278       0.02     0.548    0.018     0.030    0.047   0.007
8           0.014  295       0.02     0.533    0.008     0.017    0.029   0.005
9           0.029  347       0.02     0.492    0.014     0.024    0.038   0.006
10          0.037  349       0.02     0.491    0.017     0.029    0.043   0.007
11          0.039  358       0.02     0.484    0.018     0.030    0.045   0.007
12          0.018  396       0.02     0.459    0.010     0.019    0.030   0.005
13          0.028  431       0.02     0.438    0.015     0.024    0.037   0.006
14          0.025  441       0.02     0.433    0.013     0.023    0.035   0.005
15          0.027  477       0.02     0.414    0.015     0.024    0.036   0.006
16          0.045  484       0.02     0.410    0.023     0.035    0.050   0.007
17          0.030  494       0.02     0.405    0.016     0.026    0.039   0.006
18          0.022  501       0.02     0.402    0.012     0.021    0.032   0.005
19          0.028  505       0.02     0.400    0.015     0.025    0.036   0.005
20          0.020  540       0.02     0.384    0.012     0.020    0.031   0.005
21          0.028  563       0.02     0.374    0.016     0.025    0.037   0.005
22          0.024  593       0.02     0.362    0.014     0.022    0.033   0.005
23          0.015  602       0.02     0.358    0.010     0.017    0.026   0.004
24          0.024  629       0.02     0.348    0.014     0.023    0.033   0.005
25          0.020  636       0.02     0.346    0.012     0.020    0.030   0.005
26          0.048  729       0.02     0.316    0.027     0.039    0.053   0.007
27          0.031  849       0.02     0.284    0.019     0.028    0.038   0.005
28          0.027  914       0.02     0.269    0.017     0.025    0.035   0.005
29          0.021  940       0.02     0.264    0.014     0.021    0.030   0.004
30          0.029 1193       0.02     0.220    0.020     0.027    0.036   0.004
31          0.020 1340       0.02     0.201    0.014     0.020    0.027   0.003
colMeans    0.029  517       0.02     0.438    0.015     0.025    0.037   0.006
\end{CodeOutput}
\end{CodeChunk}
For reference, we need to type `\code{R> print(p, sort = FALSE)}' instead of `\code{R> p}' in order to list hospitals by the order of data input in the above output. `\code{R> p}' automatically sorts the output by the increasing order of $n_{j}$. 
\\

@@

As we can see in (6), the posterior mean ($(1-B_{j})y_{j} + B_{j}\lambda_{0}$) is a linear function of shrinkage ($\equiv r / (r + n_{j})$), locating between the sample mean and prior mean ($\lambda_{0}=0.02$). It makes sense because $r$ can be interpreted as the amount of prior information and $n_{j}$ as the amount of observed information. If the second level has more information than the first level, then the sample mean shrinks towards the prior mean more than 50\%. This point is clear in the above output; as caseload increases, shrinkage decreases, depending less on the prior information.
\\

@@

We also need a graphical summary that can give us valuable insight buried in pile of numbers and a function `\code{plot}' is exactly for this purpose.
\begin{CodeChunk}
\begin{CodeInput}
R> plot(p)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.3]{hospital1.png}
\end{center}
\end{figure}
The regression towards the mean (RTTM) is obvious in the left-side graph; the observed sample means are shrinking towards the known second-level mean (a blue vertical line at 0.02) to the different extents. Note that some hospitals' ranks have changed by shrinking much harder towards 0.02 than others. For example, the empty square at the crossing point of the two left-most lines (8th and 23rd hospitals on the list above) indicates that seemingly safest hospital among 31 hospitals in terms of the observed sample mean was not actually safer than the second safest hospital. Without this hierarchical modeling, we might have made a wrong decision in choosing a hospital.
\\

Intuitively, the result of multi-level modeling makes more sense than that of naive sample mean. For example, suppose there are two hospitals, whose sample means ($z_{j} / n_{j}$) are 0 and 0.01 and caseloads ($n_{j}$) are 1 and 100 each. Do you believe that the former hospital is better than the latter and are you going to choose the former hospital? Borrowing information from state-level hierarchy seems reasonable for the former hospital because it is hard to judge their true death rate per exposure with just one caseload. Though somewhat extreme, this is what happens for the two left-most hostpitals on the first plot and this is why hierarchical modeling is a reasonable choice for this dataset.
\\

@@

Also it reveals that the 23rd hospital whose estimated true rate was the smallest has also the smallest upper bound. If you are a risk-avoider, this hospital will attract you most strongly. And if you already chose this 23rd hospital compared to the 8th from the shrinkage plot, your decision might become stronger at this point, excluding the 8th hospital with more certainty. 
\\

Then, how reliable are these intervals? Does our procedure to generate interval estimates have good repeated sampling property? The method check below will answer this question.
\begin{CodeChunk}
\begin{CodeInput}
R> pcv <- coverage(p, nsim = 10000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 5.5cm]{hospital3.png}
\end{center}
\end{figure}
This \code{coverage} function generated 10,000 pseudo-datasets assuming the estimated $r~(=336.37)$ as a given value. For reference, we can try any other value of $r$, for example $r=200$, by replacing above code with `\code{R> pcv <- coverage(p, A.or.r = 200, mean.PriorDist = 0.02, nsim = 10000)}.
\\

It estimated coverage probabilities by unbiased estimates (a blue dashed line) and Rao-Blackwellized estimates (a red line). Both lines are indistingishable from the horizontal line at 0.95 and the estimated overall average coverage rate is 0.951 based on both estimation methods. Finally, this method check showed that the interval estimates from the suggested multi-level modeling on this particular dataset has a very good repeated sampling property, achieving pre-specified 95\% confidence level.

\subsection[Unknown Second-level Mean and No Covariate]{8 Schools: Unknown Second-level Mean and No Covariate}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.3]{school1.png}
\caption{Shrinkage and interval plots of 8 schools}
\end{center}
\end{figure}

@@

4. Morris, C. and Tang, R. (2011). ``Estimating Random Effects via Adjustment for Density Maximization.'' \emph{Statistical Science}. \textbf{26}. 271-287.
\\

5. Morris, C. and Lysy, M. (2012). ``Shrinkage Estimation in Multilevel Normal Models.'' \emph{Statistical Science}. \textbf{27}. 115-134.
\\

6. Rubin, D. B. (1981). ``Estimation in parallel randomized
  experiments.'' \emph{Journal of Educational Statistics}. \textbf{6}. 377-401.
\\

>>>>

Considering the type of data, it makes sense to assume the number of deaths in each hospital has independent Poisson distribution given the unknown true rate parameter $\lambda_{j}$: $z_{j}\vert \lambda_{j}\stackrel{ind}{\sim} \textrm{Poisson}(n_{j}\lambda_{j})$, $j=1, \ldots, 31,$ where $n_{j}$ can be interpreted as an exposure (not necessarily an interger) and $\lambda_{j}$ as a true death rate per exposure. This dataset is also included in the package and can be called by `\code{R> data(hospital)}' on \proglang{R}.

@@

where $\log(\frac{p_{0j}}{1-p_{0j}}) =x_{j}'\beta~\textrm{and}~j=1, \ldots, 18$. Then, corresponding posterior distribution is
\begin{equation}
p_{j}\vert \textbf{z}, \beta, r \stackrel{ind}{\sim}\textrm{Beta}(rp_{0j}+n_{j}y_{j},~r(1-p_{0j})+n_{j}(1-y_{j}))\sim\textrm{Beta}\bigg[p^{\ast}_{j},~ \frac{p^{\ast}_{j}(1-p^{\ast}_{j})}{r+n_{j}+1}\bigg],
\end{equation}
where $p^{\ast}_{j}\equiv(1-B_{j})y_{j}+B_{j}p_{0j}$, $B_{j}\equiv r/ (r+n_{j})$, and $y_{j}\equiv z_{j} / n_{j}$, $j=1,\ldots,18$.

@@

where $m$ is the number of regression coefficients. For $\beta$, it is reasonable choice to take flat (non-informative) distribution because information about the location gets plentiful as the number of groups increases. Another flat prior on the second-level variance component  $A$ (or $1/r$) that we suggest here will guarantee a posterior propriety under moderate conditions and will bring a good repeated sampling property.
\\

@@

Once we estimate these two moments of shrinkage, we can also estimate the posterior moments given only data. Taking the Normal model as an example,  $E(\mu_{j}\vert \textbf{y})=E(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})$ by Adam's law and $Var(\mu_{j}\vert \textbf{y})=E(Var(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})+Var(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})$ by Eve's law. Note that both $E(\mu_{j}\vert \textbf{y}, r, A)$ and $Var(\mu_{j}\vert \textbf{y}, r, A)$ are a function of shrinkage in (3) and Adam's and Eve's laws are taking conditional expectation and variance over them given data.

@@

\subsection{Pseudo-data Generation Process}
Figure 1 will be helpful to understand this process. As we can see in (8), the distribution of each true batting average ($p_{j},~j=1,\ldots, 18$) depends on two hyper-parameters, $r$ and $\beta_{(m\times1)}$, where $m$ is the number of regression coefficients. So, once we fix these hyper-parameters at specific values, we can generate true batting averages. Suppose we sampled 500 $\mathbf{p}_{(18\times1)}$'s, \emph{i.e.}, $\{\mathbf{p}^{(i)}_{(18\times1)},~i=1, \ldots, 500\}$ from the prior distribution in (8), where $r$ and $\beta$ are given. Then, we can also generate $\{\mathbf{z}^{(i)}_{(18\times1)},~i=1, \ldots, 500\}$ given each $p_{ij},$ where $i$ indicates $i$-th pseudo-dataset and $j$ does $j$-th player. Next, \code{coverage} fits the Binomial hierarchical model 500 times on $\{(\mathbf{z}^{(i)}_{(18\times1)}, \mathbf{n}_{(18 \times 1)}),~i=1, \ldots, 500\}$ to obtain 500$\times$18 interval estimates.  After this generating process, as for the first player, we will have 500 pairs of ($z_{i1}$, $p_{i1}$) and 500 interval estimates, $\{(\hat{p}_{i1, low}, ~\hat{p}_{i1, upp}$), $i=1, \ldots, 500\}$.
\begin{figure}[h]
\begin{center}
\includegraphics[width=5cm]{process.png}
\caption{Pseudo-data generating process}
\end{center}
\end{figure}

@@

Assuming an additional hierarchy is reasonable, a model-fitting process begins.  Since the true death rate per exposure after CABG surgery might be small and the caseload ($n_{j}$) looks relatively much bigger than the number of deaths ($z_{j}$), the Poisson distribution would be our first choice to describe the uncertainty in our data. Next, \code{gbp} will help us fit the Poisson multi-level model with the Gamma conjugate prior distribution on the true death rate $\lambda_{j}$ whose mean is 0.02 ($\lambda_{0}=0.02$) as described in 3.2 . For reference, the number of regression coefficients ($m$) is 0 because we do not need to estimate the prior mean via any regression.
\begin{CodeChunk}
\begin{CodeInput}
R> p <- gbp(z, n, mean.PriorDist = 0.02, model = "pr")
R> p
\end{CodeInput}
\begin{CodeOutput}
Summary for whole observations:

         obs.mean    n prior.mean shrinkage low.intv post.mean upp.intv post.sd
1           0.045   67       0.02     0.834    0.011     0.024    0.042   0.008
2           0.029   68       0.02     0.832    0.010     0.022    0.038   0.007
3           0.024  210       0.02     0.616    0.011     0.021    0.035   0.006
4           0.043  256       0.02     0.568    0.017     0.030    0.046   0.008
5           0.033  269       0.02     0.556    0.015     0.026    0.041   0.007
6           0.044  274       0.02     0.551    0.018     0.031    0.047   0.008
7           0.043  278       0.02     0.548    0.018     0.030    0.047   0.007
8           0.014  295       0.02     0.533    0.008     0.017    0.029   0.005
9           0.029  347       0.02     0.492    0.014     0.024    0.038   0.006
10          0.037  349       0.02     0.491    0.017     0.029    0.043   0.007
11          0.039  358       0.02     0.484    0.018     0.030    0.045   0.007
12          0.018  396       0.02     0.459    0.010     0.019    0.030   0.005
13          0.028  431       0.02     0.438    0.015     0.024    0.037   0.006
14          0.025  441       0.02     0.433    0.013     0.023    0.035   0.005
15          0.027  477       0.02     0.414    0.015     0.024    0.036   0.006
16          0.045  484       0.02     0.410    0.023     0.035    0.050   0.007
17          0.030  494       0.02     0.405    0.016     0.026    0.039   0.006
18          0.022  501       0.02     0.402    0.012     0.021    0.032   0.005
19          0.028  505       0.02     0.400    0.015     0.025    0.036   0.005
20          0.020  540       0.02     0.384    0.012     0.020    0.031   0.005
21          0.028  563       0.02     0.374    0.016     0.025    0.037   0.005
22          0.024  593       0.02     0.362    0.014     0.022    0.033   0.005
23          0.015  602       0.02     0.358    0.010     0.017    0.026   0.004
24          0.024  629       0.02     0.348    0.014     0.023    0.033   0.005
25          0.020  636       0.02     0.346    0.012     0.020    0.030   0.005
26          0.048  729       0.02     0.316    0.027     0.039    0.053   0.007
27          0.031  849       0.02     0.284    0.019     0.028    0.038   0.005
28          0.027  914       0.02     0.269    0.017     0.025    0.035   0.005
29          0.021  940       0.02     0.264    0.014     0.021    0.030   0.004
30          0.029 1193       0.02     0.220    0.020     0.027    0.036   0.004
31          0.020 1340       0.02     0.201    0.014     0.020    0.027   0.003
colMeans    0.029  517       0.02     0.438    0.015     0.025    0.037   0.006
\end{CodeOutput}
\end{CodeChunk}
For reference, we need to type `\code{R> print(p, sort = FALSE)}' instead of `\code{R> p}' in order to list hospitals by the order of data input in the above output. `\code{R> p}' automatically sorts the output by the increasing order of caseload ($n_{j}$). 
\\

@@

As we can see in (6), the posterior mean, $(1-B_{j})y_{j} + B_{j}\lambda_{0}$, is a linear function of shrinkage, $B_{j}\equiv r / (r + n_{j})$, locating between the sample mean and prior mean ($\lambda_{0}=0.02$). It makes sense because $r$ can be interpreted as the amount of prior information and $n_{j}$ as the amount of observed information. If the second level has more information than the first level, then the sample mean shrinks towards the prior mean more than 50\%. This point is clear in the above output; as caseload increases, shrinkage decreases, depending less on the prior information.
\\

@@

We also need a graphical summary that can give us valuable insight buried in pile of numbers and a function `\code{plot}' is exactly for this purpose.
\begin{CodeChunk}
\begin{CodeInput}
R> plot(p)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.3]{hospital1.png}
\end{center}
\end{figure}
The regression towards the mean (RTTM) is obvious in the left-side graph; the observed sample means, empty dots on the upper line, are shrinking towards the known second-level mean (a blue vertical line at 0.02) to the different extents. Note that some hospitals' ranks have changed by shrinking much harder towards 0.02 than others. For example, the empty square symbol at the crossing point of the two left-most lines (8th and 23rd hospitals on the list above) indicates that seemingly safest hospital among 31 hospitals in terms of the observed sample mean was not actually safer than the second safest hospital. Without this hierarchical modeling, we might have made a wrong decision in choosing a hospital.
\\

Intuitively, the result of multi-level modeling makes more sense than that of naive sample mean. For example, suppose there are two hospitals, whose sample means ($z_{j} / n_{j}$) are 0 and 0.01 and caseloads ($n_{j}$) are 1 and 100 each. Do you believe that the former hospital is better than the latter and are you going to choose the former hospital? Borrowing information from state-level hierarchy seems reasonable for the former hospital because it is hard to judge its true death rate per exposure with just one caseload. Though somewhat extreme, this is what happens for the two left-most hostpitals on the first plot and this is why hierarchical modeling is a reasonable choice for this dataset.
\\

@@

Also it reveals that the 23rd hospital, whose estimated true rate was the smallest, has also the smallest upper bound. If you are a risk-avoider, this hospital will attract you most strongly. And if you already chose this 23rd hospital compared to the 8th from the shrinkage plot, your decision might become stronger at this point, excluding the 8th hospital with more certainty. 
\\

Then, how reliable are these intervals? Does our procedure to generate interval estimates have good repeated sampling property? The method check below will answer this question.
\begin{CodeChunk}
\begin{CodeInput}
R> pcv <- coverage(p, nsim = 10000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 5.5cm]{hospital2.png}
\end{center}
\end{figure}
This \code{coverage} function generated 10,000 pseudo-datasets assuming the estimated $r~(=336.37)$ as a given value. For reference, we can try any other value of $r$, for example $r=200$, by replacing above code with `\code{R> pcv <- coverage(p, A.or.r = 200, mean.PriorDist = 0.02, nsim = 10000)}.
\\

It estimated coverage probabilities by unbiased estimates (a blue dashed line) and Rao-Blackwellized estimates (a red line). Both lines are indistingishable from the horizontal line at 0.95 and the estimated overall average coverage rate is 0.951 based on both estimation methods. Finally, this method check showed that the interval estimates from the suggested multi-level modeling on this particular dataset has a very good repeated sampling property, achieving pre-specified 95\% confidence level.
\\

Morris and Christiansen (1995) also look into this ranking problem, taking shrinkage into account.
\subsection[Unknown Second-level Mean and No Covariate]{8 Schools: Unknown Second-level Mean and No Covariate}
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.3]{school1.png}
\caption{Shrinkage and interval plots of 8 schools}
\end{center}
\end{figure}

@@

4. Morris, C. and Christiansen, C. (1995). ``Hierarchical Models for Ranking and for Identifying Ex- tremes, With Application,'' in \emph{Bayesian Statistics 5}, eds. J. Bernardo, J. Berger, A. Dawid, and A. Smith, New York: Oxford University Press, pp. 277-296.
\\

5. Morris, C. and Tang, R. (2011). ``Estimating Random Effects via Adjustment for Density Maximization.'' \emph{Statistical Science}. \textbf{26}. 271-287.
\\

6. Morris, C. and Lysy, M. (2012). ``Shrinkage Estimation in Multilevel Normal Models.'' \emph{Statistical Science}. \textbf{27}. 115-134.
\\

7. Rubin, D. B. (1981). ``Estimation in parallel randomized
  experiments.'' \emph{Journal of Educational Statistics}. \textbf{6}. 377-401.
\\
